Left-linear rewriting in adhesive categories

### 2024-04-23 concur/fusione.tex
minima

\begin{example}
  \label{ex:seq-ind}
  Consider a rewriting system in $\cat{Graph}$, the category of
  directed graphs and graph morphisms, which is adhesive. The set of
  rules $\R = \{ \rho_0, \rho_1, \rho_2\}$ is in
  Fig.~\ref{fi:rules}, where numbers are used to represent the
  morphisms from the interface to the left- and right-hand
  sides. Rules $\rho_0$ and $\rho_1$ are linear: $\rho_0$ generates a
  new node and a new edge, while $\rho_1$ creates a self-loops edge. Instead,
  $\rho_2$ is left-linear but not linear, as it ``merges'' two nodes.

@@

  As a consequence, $\der{D}$ and $\der{D}'$ would be switch
  equivalent, but this is not acceptable when viewing equivalence
  classes of derivations as concurrent computations: in $\der{D}$ the
  first two steps are not sequential independent, while in $\der{D}'$
  they are, intuitively because in $\der{D}$ rule $\rho_1$ uses the
  node generated by $\rho_0$ (adding a self loop to it), while in
  $\der{D}'$ rule $\rho_1$ uses the node that was in the initial
  graph. Also observe that the graphs $G_2$ and $G_2'$ produced after
  two steps in $\der{D}$ and $\der{D'}$ are not isomorphic. From the
  technical point of view, the property of being switch equivalent is
  not closed by prefix, and this prevents to derive a sensible
  concurrent semantics: In fact
  $\der{D} = \dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2$ and
  $\der{D}' = \dder{D}_0'\cdot \dder{D}_1' \cdot \dder{D}_2'$ are
  switch equivalent, while if we consider the first two steps,
  derivations $\dder{D}_0 \cdot \dder{D}_1$ and
  $\dder{D}_0' \cdot \dder{D}_1'$ are not switch equivalent.
  % if we limit to the first
  %two steps, we obtain derivations $\dder{D}_0 \cdot \dder{D}_1$ and
  %$\dder{D}_0' \cdot \dder{D}_1'$ that are not switch equivalent, while
  %$\der{D} = \dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2$ and
  %$\der{D}' = \dder{D}_0'\cdot \dder{D}_1' \cdot \dder{D}_2'$ would
  %be.
\end{example}

@@

This result fails when considering left-linear
derivations. This is readily seen by considering derivations $\der{D}$
in Fig.~\ref{fi:derD} and $\der{D}'$ in Fig.~\ref{fi:derD1}. We
observed that $\der{D}$ and $\der{D}'$ are not switch equivalent.
% 
Still, it is easy to see that they are related by a consistent
permutation, which is the identity on productions. The colimit of
both derivations is a single node with a self loop, i.e.~the
terminal object in $\cat{Graph}$, hence all commutativity requirements
are trivially satisfied.

@@

\begin{example}
  \label{ex:non-global}
  Consider a new rewriting system in $\cat{Graph}$ consisting of the
  rules $\sigma_0$, $\sigma_1$, and $\sigma_2$ in
  Fig.~\ref{fi:rules-cas}.
  %\todo{P: rev.2 dice ``In the top-left and
   % top-middle nodes of $\rho_3$, the labels "12" should be "1" and
   % "2".''. Non e' corretto, ma magari si capisce poco ...}  
   Consider the derivation $\der{F}$ in Fig.~\ref{fi:derF}. Note that rule
  $\sigma_2$ needs the black self-loop on node $12$ to be applied. This can be
  generated either by $\sigma_0$ or by $\sigma_1$ merging nodes $1$ and
  $2$, hence at least one of them must precede the application of
  $\sigma_2$. Indeed, in $\der{F}$ it can be easily seen that the second
  and the third step applying $\sigma_1$ and $\sigma_2$ are
  independent. However, we could switch twice the application of
  $\sigma_0$, bringing it in the last position, obtaining a derivation
  $\der{F}'$ as depicted in Fig.~\ref{fi:derF1}, where the
  applications of $\sigma_1$ and $\sigma_2$ are no longer independent.

@@

While inversions can be
performed in any order in the case of linear rewriting systems, this is no longer true for left-linear
rewriting systems due to the fact that, as already observed, globality of independence holds in a weak form. 
For instance, if we get back to Example~\ref{ex:non-global} and consider derivation $\der{F}$ in Fig.~\ref{fi:derF}, 
the last two steps applying rules $\sigma_1$ and $\sigma_2$ can be switched, thus leading to a derivation, 
call it $\der{F}''$,  applying $\sigma_0, \sigma_2, \sigma_1$. Such $\der{F}''$ is switch equivalent to $\der{F}'$ in Fig.~\ref{fi:derF1}, 
hence starting from $\der{F}''$ we can obtain $\der{F}'$ via a switching sequence, but we cannot start 
by switching $\sigma_0$ and $\sigma_2$, despite the fact that this is an inversion.

@@

The previous proposition, now, allows us to establish the following result.

@@

We are now going to apply the results of the previous section to $\mathcal{M}$-adhesive categories in order to establish some \emph{high-level replacement properties} \cite{ehrig2004adhesive,ehrig2014adhesive,ehrig2006fundamentals}.  
A first important  result which can be immediately established, with the aid of \Cref{prop:pbpo}, is the following one.

@@

	The thesis now follows immediately from \Cref{lem:varie}. \qedhere
\end{proof}
}
\noindent
\parbox{10cm}{
	\begin{corollary}[Uniqueness of pushouts complements]\label{lem:pocomp}
		Let $\X$ be a $\mathcal{M}$-adhesive category. Given $m\colon X\mto Y$ in $\mathcal{M}$ and $f\colon Y\to Z$, let $(h_1, k_1)$ and $(h_2, k_2)$ be respectively the pushout complements of $m$  and $f$ depicted aside. Then there exists a unique isomorphism $\phi\colon W_1\to W_2$ making the diagram on the right commutative.
	\end{corollary}}
\parbox{3cm}{\vspace{1em}$\xymatrix{&X \ar@{>->}[r]^{m} \ar[d]_{h_1} \ar@/_.3cm/[dl]_{h_2}& Y \ar[d]^{f} \\ W_2 \ar@{>->}@/_.4cm/[rr]_{k_2} &W_1 \ar@{.>}[l]_{\phi} \ar@{>->}[r]^{k_1} & Z }$}

@@

\section{$\mathcal{M}$-adhesivity is not enough}\label{app:fill}
In \Cref{sec:ade} we introduced the notion of strong enforcing left-linear rewriting system, a class that contains, for instance, all the linear rewriting systems. This result can be further refined: in \cite{baldan2011adhesivity} a class  $\mathbb{B}$ of (quasi-)adhesive category is defined for which the local Church-Rosser Theorem holds for left-linear rewriting system. In our language, this means that every left-linear rewriting system based on a category in $\mathbb{B}$ is strong enforcing. This section is devoted to repropose, and slightly generalise, the results of \cite{baldan2011adhesivity} to our context.

@@

\noindent
\parbox{10.7cm}{\begin{proposition}\label{prop:po2}
		Let $(\X, \mathcal{M})$ be a pair of type $\mathbb{B}$, and suppose that the square aside is an $\mathcal{M}$ pullback. Then there exists $E\in \X$, $e\colon E\mto C$ in $\mathcal{M}$, $\phi:E\to B_1$ such that $(C, \{m, e\})$ is a coproduct and $g=f+\phi$. Moreover, such a square is a pushout if and only if $\phi$ is an isomorphism.
\end{proposition}} \parbox{3cm}{\vspace{-.2cm}$\xymatrix{A  \ar[r]^{f} \ar@{>->}[d]_{m}& B_0 \ar@{>->}[d]^{\iota_{B_{0}}}\\ C\ar[r]_-{g} & B_0+B_1}$}
\iffalse 
\begin{proof}
		Pulling back $\iota_{B_0}\colon B_0\to B_0\to B_1$ and $\iota_{B_1}\colon B_1\to B_0 + B_1$ we get the diagram below and the thesis follows from \Cref{prop:ext}. The same diagram allows us to deduce $g=f+\phi$.	
	\parbox{5.2cm}{$\xymatrix{A\ar[r]^-{m} \ar@{>->}[d]_{r}& C  \ar@{>->}[d]_{g}& E \ar@{>->}[d]^{\phi}\ar[l]_-{e}\\B_0  \ar[r]_-{\iota_{B_0}}& B_0+B_1 & B_1 \ar[l]^-{\iota_{B_1}} \\ A & B_0 \\C & B_0+E\\ &&B_0+B_1}$}\parbox{8.8cm}{\vspace{-4.1cm}Let us show the second half of the thesis.

@@

	\medskip \noindent
	\parbox{3cm}{$\xymatrix{X \ar[r]^f \ar@{>->}[d]_{\iota_X}& Y\ar[r]^g \ar@{>->}[d]^{\iota_Y} & Z\ar@{>->}[d]^{\iota_Z}\\X+A  \ar@/_.4cm/[rr]_{(g\circ f) + (\varphi \circ \phi)}\ar[r]^{f+\phi}& Y+B \ar[r]^{g+\varphi}& Z+C}$} \qquad \qquad  \qquad   \qquad \parbox{8cm}{ $\mathcal{M}$-pushout-decomposition property. Using \Cref{prop:ext,prop:po2} and the fact that arrows in $\mathcal{M}$ are coproduct coprojections, we can reduce to prove the property for diagrams as the one on the left. By hypothesis and \Cref{prop:po2}, $\varphi$ and $\varphi\circ \phi$ are both isomorphisms. therefore $\phi$ is an isomorphism too and we can conclude again using \Cref{prop:po2}.\qedhere 
}

@@

\begin{proposition}
  \label{prop:unique}
  Let $\X$ be a $\mathcal{M}$-adhesive category. Suppose that the two
  direct derivations $\dder{D}$ and $\dder{D}'$ below, with the same
  match and applying the same left-linear rule $\rho$, are given
  \[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} &
      R \ar[d]^{h} & L \ar[d]_{m}& K \ar[d]^{k'}\ar@{>->}[l]_{l}
      \ar[r]^{r} & R \ar[d]^{h'}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H
      & G & \ar@{>->}[l]^{f'} D' \ar[r]_{g'}& H'}\]
  Then there is an abstraction equivalence between $\dder{D}$ and
  $\dder{D}'$, whose first component is $\id{G}$.
\end{proposition}
\begin{proof}
  Both pairs $(k, f)$ and $(k', f')$ are pushout complements of $l$
  and $n$, thus, \Cref{lem:pocomp} yields an isomorphism
  $\phi_D\colon D\to D'$. An easy computation shows that $g'\circ \phi_D \circ k=h'\circ r$.
  \noindent
  \parbox{3cm}{
    $\xymatrix{K \ar@/^.4cm/[rr]^{k'}\ar[d]_{r} \ar[r]_{k}&
      \ar[r]_{\phi_D} D \ar[d]^{g}& D' \ar[d]^{g'}\\ R
      \ar@/_.4cm/[rr]_{h'} \ar[r]^{h}& H \ar[r]^{\phi_H}& H'}$}
  \hfill
  \parbox{10cm}{ \hspace{15pt}Hence, we have another
    $\phi_H\colon H\to H'$. To see that $\phi_H$ is an isomorphism,
    consider the diagram aside. By hypothesis the whole rectangle and
    its left half are pushouts, therefore, by \Cref{lem:po1} its right
    square is a pushout too. The claim now follows from the fact that
    the pushout of an isomorphism is an isomorphism. \qedhere }
\end{proof}

@@

There is another
useful property of  sequentially independent direct
derivations which is worth mentioning.

@@

\begin{proof}
  Consider the arrow $u_1\colon K_1\to P$ obtained using
  \Cref{prop:tec}. It fits in the big square on the left:
\[		\xymatrix@R=15pt{K_1 \ar[r]^{u_1} \ar@{>->}[d]_{l_1}&P
	\ar[r]^{\id{P}}\ar@{>->}[d]_{p_0}& P\ar@{>->}[d]^{p_0}\\L_1
	\ar[d]_{\id{L_1}}\ar[r]^{i_1}&D_0 \ar[r]^{\id{D_0}}
	\ar[d]_{\id{D_0}}& D_{0} \ar@{>->}[d]^{f_0}&K_1 \ar@{>->}[d]_{l_1}\ar[r]^{k_{\der{E},0}}&
	D_{\der{E},0} \ar@{>->}[d]^{f_{\der{E},0}} & K_1 \ar[r]^{u}
	\ar@{>->}[d]_{l_1}& P \ar@{>->}[d]^{f_0\circ
		p_0}\\
	L_1\ar[r]^{i_1} \ar@/_.4cm/[rr]_{m_{\der{E},0}}&D_0
	\ar@{>->}[r]^{f_0}&G_0 & L_1\ar[r]_{m_{\der{E},0}}&G_0 &L_1
	\ar[r]_{m_{\der{E},0}}&G_0}\]

@@

\begin{remark}
	\label{rem:church}
	It is worth to point out that, by
	\Cref{prop:switch,thm:switch_uni}, the derivation we obtain from
	\Cref{thm:church} is made by two switchable derivations that will
	give back the starting derivation (up to abstraction equivalence),
	if switched again.

@@

	The arrow $f_1\colon D\to G_1$ is the pushout of
	$l_1\colon K_1\mto L_1$, thus it is in $\mathcal{M}$ implying
	$i_0=i'_0$. If, moreover, we suppose that the rule $\rho$ applied in
	$\dder{D}$ is linear, then $g_0\colon D_0\to G_1$ is in
	$\mathcal{M}$ too entailing $i_1=i'_1$, too.
\end{proof}

@@

\begin{definition}[Consistent permutation]
	\label{def:permcon}
	Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X, \R)$ 
	a left-linear rewriting system.  Take two
	non-empty decorated derivations $(\der{D}, \alpha, \omega)$ and
	$(\der{D}', \alpha', \omega')$ with the same length and with
	isomorphic sources and targets.

@@

\begin{proof}   
	Since $(\X, \R)$ is well-switching, there exists unique independence pairs 
$(i_0,i_1)$ between $\dder{D}_0$ and $\dder{D}_1$,	$(a_0,a_1)$ between $\dder{D}_1$ and $\dder{D}_2$ and	$(e_0, e_1)$ one between $\dder{D}_0$ and
	$S_{a_0,a_1}(\dder{D}_2)$. 	Using \Cref{rem:fill} we can build the diagrams in \Cref{fi:first,fi:second,fi:third}.  The first depicts the  switching $\dder{D}_0$ and $\dder{D}_1$ along  $(i_0, i_1)$, the second that of $\dder{D}_1$ and $\dder{D}_2$ along $a_0$ and $a_1$, while the third refers to $\dder{D}_0$ and
	$S_{a_0,a_1}(\dder{D}_2)$.

@@

	Finally, switch the second and the third step we get
	\[\xymatrix@C=15pt{L_2\ar[d]^{f_0\circ z_1}&& K_2 \ar[d]_{\mathring{k}_2}\ar@{>->}[ll]_{l_2} \ar[r]^{r_2} & R_2 \ar@/^.35cm/[drrr]_(.4){v_0}|(.29)\hole \ar[dr]|(.28)\hole_{\mathring{h}_2} && L_1 \ar@/_.35cm/[dlll]^(.4){v_1} \ar[dl]|(.28)\hole^{\mathring{f}_1 \circ x_1}& K_1 \ar[d]^{\mathring{k}_1}\ar@{>->}[l]_{l_1} \ar[r]^{r_1} & R_1 \ar@/^.35cm/[drrr]_(.4){u_0}|(.29)\hole\ar[dr]_{\check{h}_1}|(.28)\hole   && L_0 \ar@/_.35cm/[dlll]^(.4){u_1} \ar[dl]^{\mathring{m}_0}|(.28)\hole& K_0 \ar[d]^{\mathring{k}_0}\ar@{>->}[l]_{l_0} \ar[rr]^{r_0} && R_0\ar[d]_{\check{g}_2\circ x_0} \\G_0 && \ar[ll]^{\check{f}_0} \check{D}_0 \ar[rr]_{\check{g}_0}&& \check{G}_1  && \ar@{>->}[ll]^{\bar{f}_1} \bar{D}_1 \ar[rr]_{\bar{g}_1}&& \bar{G}_2 && \ar@{>->}[ll]^{\bar{f}_2} \bar{D}_2 \ar[rr]_{\bar{g}_2}&& G_3 }\]
	As usual, this last switch brings identities $\check{m}_0=\bar{f}_1\circ u_1$ and $g_2\circ j_0=\bar{g}_2\circ u_0$. Applying the second point of \Cref{lem:indep-global-left} to $\dder{D}_0\cdot S_{j_0, j_1}(\dder{D}_2)\cdot S_{j_0, j_1}(\dder{D}_1)$ and the fact that $(\X, \R)$ is well-switching we get the following diagrams

@@

We next prove some technical lemma that will be useful to show the
possibility of transforming a derivation into a switch equivalent one
by switching only step along inversions.

@@

\begin{lemma}
  \label{le:switchB}
  Let $\der{D}_i$ for $i \in \interval[0]{n}$ be derivation sequences
  and consider a switching sequence consisting only of inversions
  \begin{center}
    $\der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n$
  \end{center}
  Let $i \in \interval{n}$ be such that
  $\transp{i} \in \inv{\nu_{1,n}}$ and assume that steps $i$ and $i+1$ in $\der{D}_0$ are sequential independent. Then there is a switching sequence
  consisting only of inversions which starts with $\transp{i}$, i.e.,
  \begin{center}
    $\der{D}_0 \shift{\transp{i}} \der{D}_1' \shift{\nu_2'} \ldots
    \shift{\nu_n'} \der{D}_n$
  \end{center}
\end{lemma}

>>>>

\begin{example}
  \label{ex:seq-ind}
  Consider a rewriting system in $\cat{Graph}$, the category of
  directed graphs and graph morphisms, which is adhesive. The set of
  rules $\R = \{ \rho_0, \rho_1, \rho_2\}$ is in
  Fig.~\ref{fi:rules}, where numbers are used to represent the
  morphisms from the interface to the left- and right-hand
  sides. Rules $\rho_0$ and $\rho_1$ are linear: $\rho_0$ generates a
  new node and a new edge, while $\rho_1$ creates a self-loop edge. Instead,
  $\rho_2$ is left-linear but not linear, as it ``merges'' two nodes.

@@

  As a consequence, $\der{D}$ and $\der{D}'$ would be switch
  equivalent, but this is not acceptable when viewing equivalence
  classes of derivations as concurrent computations: in $\der{D}$ the
  first two steps are not sequential independent, while in $\der{D}'$
  they are, intuitively because in $\der{D}$ rule $\rho_1$ uses the
  node generated by $\rho_0$ (adding a self-loop to it), while in
  $\der{D}'$ rule $\rho_1$ uses the node that was in the initial
  graph. Also observe that the graphs $G_2$ and $G_2'$ produced after
  two steps in $\der{D}$ and $\der{D'}$ are not isomorphic. From the
  technical point of view, the property of being switch equivalent is
  not closed by prefix, and this prevents to derive a sensible
  concurrent semantics: In fact
  $\der{D} = \dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2$ and
  $\der{D}' = \dder{D}_0'\cdot \dder{D}_1' \cdot \dder{D}_2'$ are
  switch equivalent, while if we consider the first two steps,
  derivations $\dder{D}_0 \cdot \dder{D}_1$ and
  $\dder{D}_0' \cdot \dder{D}_1'$ are not switch equivalent.
  % if we limit to the first
  %two steps, we obtain derivations $\dder{D}_0 \cdot \dder{D}_1$ and
  %$\dder{D}_0' \cdot \dder{D}_1'$ that are not switch equivalent, while
  %$\der{D} = \dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2$ and
  %$\der{D}' = \dder{D}_0'\cdot \dder{D}_1' \cdot \dder{D}_2'$ would
  %be.
\end{example}

@@

This result fails when considering left-linear
derivations. This is readily seen by considering derivations $\der{D}$
in Fig.~\ref{fi:derD} and $\der{D}'$ in Fig.~\ref{fi:derD1}. We
observed that $\der{D}$ and $\der{D}'$ are not switch equivalent.
% 
Still, it is easy to see that they are related by a consistent\todo{le definiamo solo nell'appendice}
permutation, which is the identity on productions. The colimit of
both derivations is a single node with a self-loop, i.e.~the
terminal object in $\cat{Graph}$, hence all commutativity requirements
are trivially satisfied.

@@

\begin{example}
  \label{ex:non-global}
  Consider the rewriting system in $\cat{Graph}$ consisting of the
  rules $\lambda_0$, $\lambda_1$, and $\lambda_2$ in
  Fig.~\ref{fi:rules-cas}.
  %\todo{P: rev.2 dice ``In the top-left and
   % top-middle nodes of $\rho_3$, the labels "12" should be "1" and
   % "2".''. Non e' corretto, ma magari si capisce poco ...}  
   Consider the derivation $\der{F}$ in Fig.~\ref{fi:derF}. Note that rule
  $\lambda_2$ needs the black self-loop on node $12$ to be applied. This can be
  generated either by $\lambda_0$ or by $\lambda_1$ merging nodes $1$ and
  $2$, hence at least one of them must precede the application of
  $\lambda_2$. Indeed, in $\der{F}$ it can be easily seen that the second
  and the third step applying $\lambda_1$ and $\lambda_2$ are
  independent. However, we could switch twice the application of
  $\lambda_0$, bringing it in the last position, obtaining a derivation
  $\der{F}'$ as depicted in Fig.~\ref{fi:derF1}, where the
  applications of $\lambda_1$ and $\lambda_2$ are no longer independent.

@@

While inversions can be
performed in any order in the case of linear rewriting systems, this is no longer true for left-linear
rewriting systems due to the fact that, as already observed, globality of independence holds in a weak form. 
For instance, if we get back to Example~\ref{ex:non-global} and consider derivation $\der{F}$ in Fig.~\ref{fi:derF}, 
the last two steps applying rules $\sigma_1$ and $\sigma_2$ can be switched, thus leading to a derivation, 
call it $\der{F}''$,  applying $\lambda_0, \lambda_2, \lambda_1$. Such $\der{F}''$ is switch equivalent to $\der{F}'$ in Fig.~\ref{fi:derF1}, 
hence starting from $\der{F}''$ we can obtain $\der{F}'$ via a switching sequence, but we cannot start 
by switching $\lambda_0$ and $\lambda_2$, despite the fact that this is an inversion.

@@

The previous proposition allows us to establish the following result.

@@

We are now going to apply the results of the previous section to $\mathcal{M}$-adhesive categories in order to establish some \emph{high-level replacement properties} \cite{ehrig2004adhesive,ehrig2014adhesive,ehrig2006fundamentals}.  
A first important  result that can be immediately established, with the aid of \Cref{prop:pbpo}, is the following one.

@@

	The thesis now follows immediately from \Cref{lem:varie}. \qedhere
\end{proof}
}
\noindent
\parbox{10cm}{
	\begin{corollary}[Uniqueness of pushout complements]\label{lem:pocomp}
		Let $\X$ be a $\mathcal{M}$-adhesive category. Given $m\colon X\mto Y$ in $\mathcal{M}$ and $f\colon Y\to Z$, let $(h_1, k_1)$ and $(h_2, k_2)$ be respectively the pushout complements of $m$  and $f$ depicted aside. Then there exists a unique isomorphism $\phi\colon W_1\to W_2$ making the diagram on the right commutative.
	\end{corollary}}
\parbox{3cm}{\vspace{1em}$\xymatrix{&X \ar@{>->}[r]^{m} \ar[d]_{h_1} \ar@/_.3cm/[dl]_{h_2}& Y \ar[d]^{f} \\ W_2 \ar@{>->}@/_.4cm/[rr]_{k_2} &W_1 \ar@{.>}[l]_{\phi} \ar@{>->}[r]^{k_1} & Z }$}

@@

\section{$\mathcal{M}$-adhesivity is not enough}\label{app:fill}
In \Cref{sec:ade} we introduced the notion of strong enforcing left-linear rewriting system, a class that contains all the linear rewriting systems. This result can be further refined: in \cite{baldan2011adhesivity} a class  $\mathbb{B}$ of (quasi-)adhesive category is defined for which the local Church-Rosser Theorem holds for left-linear rewriting system. In our language, this means that every left-linear rewriting system based on a category in $\mathbb{B}$ is strong enforcing. This section is devoted to repropose, and slightly generalise, the results of \cite{baldan2011adhesivity} to our context.

@@

\noindent
\parbox{10.7cm}{\begin{proposition}\label{prop:po2}
		Let $(\X, \mathcal{M})$ be a pair of type $\mathbb{B}$, and suppose that the square aside is an $\mathcal{M}$-pullback. Then there exists $E\in \X$, $e\colon E\mto C$ in $\mathcal{M}$, $\phi:E\to B_1$ such that $(C, \{m, e\})$ is a coproduct and $g=f+\phi$. Moreover, such a square is a pushout if and only if $\phi$ is an isomorphism.
\end{proposition}} \parbox{3cm}{\vspace{-.2cm}$\xymatrix{A  \ar[r]^{f} \ar@{>->}[d]_{m}& B_0 \ar@{>->}[d]^{\iota_{B_{0}}}\\ C\ar[r]_-{g} & B_0+B_1}$}
\iffalse 
\begin{proof}
		Pulling back $\iota_{B_0}\colon B_0\to B_0\to B_1$ and $\iota_{B_1}\colon B_1\to B_0 + B_1$ we get the diagram below and the thesis follows from \Cref{prop:ext}. The same diagram allows us to deduce $g=f+\phi$.	
	\parbox{5.2cm}{$\xymatrix{A\ar[r]^-{m} \ar@{>->}[d]_{r}& C  \ar@{>->}[d]_{g}& E \ar@{>->}[d]^{\phi}\ar[l]_-{e}\\B_0  \ar[r]_-{\iota_{B_0}}& B_0+B_1 & B_1 \ar[l]^-{\iota_{B_1}} \\ A & B_0 \\C & B_0+E\\ &&B_0+B_1}$}\parbox{8.8cm}{\vspace{-4.1cm}Let us show the second half of the thesis.

@@

	\medskip \noindent
	\parbox{3cm}{$\xymatrix{X \ar[r]^f \ar@{>->}[d]_{\iota_X}& Y\ar[r]^g \ar@{>->}[d]^{\iota_Y} & Z\ar@{>->}[d]^{\iota_Z}\\X+A  \ar@/_.4cm/[rr]_{(g\circ f) + (\varphi \circ \phi)}\ar[r]^{f+\phi}& Y+B \ar[r]^{g+\varphi}& Z+C}$} \qquad \qquad  \qquad   \qquad \parbox{8cm}{ $\mathcal{M}$-pushout-decomposition property. Using \Cref{prop:ext,prop:po2} and the fact that arrows in $\mathcal{M}$ are coproduct coprojections, we can reduce to prove the property for diagrams as the one on the left. By hypothesis and \Cref{prop:po2}, $\varphi$ and $\varphi\circ \phi$ are both isomorphisms, therefore $\phi$ is an isomorphism too and we can conclude again using \Cref{prop:po2}.\qedhere 
}

@@

\begin{proposition}
  \label{prop:unique}
  Let $\X$ be an $\mathcal{M}$-adhesive category. Suppose that the two
  direct derivations $\dder{D}$ and $\dder{D}'$ below, with the same
  match and applying the same left-linear rule $\rho$, are given
  \[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} &
      R \ar[d]^{h} & L \ar[d]_{m}& K \ar[d]^{k'}\ar@{>->}[l]_{l}
      \ar[r]^{r} & R \ar[d]^{h'}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H
      & G & \ar@{>->}[l]^{f'} D' \ar[r]_{g'}& H'}\]
  Then there is an abstraction equivalence between $\dder{D}$ and
  $\dder{D}'$, whose first component is $\id{G}$.
\end{proposition}
\begin{proof}
  Both pairs $(k, f)$ and $(k', f')$ are pushout complements of $l$
  and $n$, thus, \Cref{lem:pocomp} yields an isomorphism
  $\phi_D\colon D\to D'$. An easy computation shows that $g'\circ \phi_D \circ k=h'\circ r$.
  \noindent
  \parbox{3cm}{
    $\xymatrix{K \ar@/^.4cm/[rr]^{k'}\ar[d]_{r} \ar[r]_{k}&
      \ar[r]_{\phi_D} D \ar[d]^{g}& D' \ar[d]^{g'}\\ R
      \ar@/_.4cm/[rr]_{h'} \ar[r]^{h}& H \ar[r]^{\phi_H}& H'}$}
  \hfill
  \parbox{10cm}{ \hspace{15pt}Hence, we have another
    $\phi_H\colon H\to H'$. To see that $\phi_H$ is an isomorphism,
    consider the diagram aside. By hypothesis the whole rectangle and
    its left half are pushouts, therefore, by \Cref{lem:po1} its right
    square is a pushout too. The claim now follows from the fact that
    the pushout of an isomorphism is an isomorphism. \qedhere }
\end{proof}

@@

There is another
useful property of  sequentially independent direct
derivations that is worth mentioning.

@@

\begin{proof}
  Consider the arrow $u_1\colon K_1\to P$ obtained using
  \Cref{prop:tec}. It fits in the big square on the left
\[		\xymatrix@R=15pt{K_1 \ar[r]^{u_1} \ar@{>->}[d]_{l_1}&P
	\ar[r]^{\id{P}}\ar@{>->}[d]_{p_0}& P\ar@{>->}[d]^{p_0}\\L_1
	\ar[d]_{\id{L_1}}\ar[r]^{i_1}&D_0 \ar[r]^{\id{D_0}}
	\ar[d]_{\id{D_0}}& D_{0} \ar@{>->}[d]^{f_0}&K_1 \ar@{>->}[d]_{l_1}\ar[r]^{k_{\der{E},0}}&
	D_{\der{E},0} \ar@{>->}[d]^{f_{\der{E},0}} & K_1 \ar[r]^{u}
	\ar@{>->}[d]_{l_1}& P \ar@{>->}[d]^{f_0\circ
		p_0}\\
	L_1\ar[r]^{i_1} \ar@/_.4cm/[rr]_{m_{\der{E},0}}&D_0
	\ar@{>->}[r]^{f_0}&G_0 & L_1\ar[r]_{m_{\der{E},0}}&G_0 &L_1
	\ar[r]_{m_{\der{E},0}}&G_0}\]

@@

\begin{remark}
	\label{rem:church}
	It is worth to point out that, by
	\Cref{prop:switch,thm:switch_uni}, the derivation we obtain from
	\Cref{thm:church}\todo{missing} is made by two switchable derivations that will
	give back the starting derivation (up to abstraction equivalence),
	if switched again.

@@

	The arrow $f_1\colon D\to G_1$ is the pushout of
	$l_1\colon K_1\mto L_1$, thus it is in $\mathcal{M}$ implying
	$i_0=i'_0$. If, moreover, we suppose that the rule $\rho$ applied in
	$\dder{D}$ is linear, then $g_0\colon D_0\to G_1$ is in
	$\mathcal{M}$ too, entailing $i_1=i'_1$.
\end{proof}

@@

\begin{definition}[Consistent permutation]
	\label{def:permcon}
	Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X, \R)$ 
	a left-linear rewriting system.  Take two
	non-empty decorated\todo{never introduced} derivations $(\der{D}, \alpha, \omega)$ and
	$(\der{D}', \alpha', \omega')$ with the same length and with
	isomorphic sources and targets.

@@

\begin{proof}   
	Since $(\X, \R)$ is well-switching, there exists unique independence pairs 
$(i_0,i_1)$ between $\dder{D}_0$ and $\dder{D}_1$,	$(a_0,a_1)$ between $\dder{D}_1$ and $\dder{D}_2$ and	$(e_0, e_1)$ between $\dder{D}_0$ and
	$S_{a_0,a_1}(\dder{D}_2)$. 	Using \Cref{rem:fill} we can build the diagrams in \Cref{fi:first,fi:second,fi:third}.  The first depicts the  switching $\dder{D}_0$ and $\dder{D}_1$ along  $(i_0, i_1)$, the second that of $\dder{D}_1$ and $\dder{D}_2$ along $(a_0, a_1)$, while the third refers to $\dder{D}_0$ and
	$S_{a_0,a_1}(\dder{D}_2)$.

@@

	Finally, switching the second and the third step we get
	\[\xymatrix@C=15pt{L_2\ar[d]^{f_0\circ z_1}&& K_2 \ar[d]_{\mathring{k}_2}\ar@{>->}[ll]_{l_2} \ar[r]^{r_2} & R_2 \ar@/^.35cm/[drrr]_(.4){v_0}|(.29)\hole \ar[dr]|(.28)\hole_{\mathring{h}_2} && L_1 \ar@/_.35cm/[dlll]^(.4){v_1} \ar[dl]|(.28)\hole^{\mathring{f}_1 \circ x_1}& K_1 \ar[d]^{\mathring{k}_1}\ar@{>->}[l]_{l_1} \ar[r]^{r_1} & R_1 \ar@/^.35cm/[drrr]_(.4){u_0}|(.29)\hole\ar[dr]_{\check{h}_1}|(.28)\hole   && L_0 \ar@/_.35cm/[dlll]^(.4){u_1} \ar[dl]^{\mathring{m}_0}|(.28)\hole& K_0 \ar[d]^{\mathring{k}_0}\ar@{>->}[l]_{l_0} \ar[rr]^{r_0} && R_0\ar[d]_{\check{g}_2\circ x_0} \\G_0 && \ar[ll]^{\check{f}_0} \check{D}_0 \ar[rr]_{\check{g}_0}&& \check{G}_1  && \ar@{>->}[ll]^{\bar{f}_1} \bar{D}_1 \ar[rr]_{\bar{g}_1}&& \bar{G}_2 && \ar@{>->}[ll]^{\bar{f}_2} \bar{D}_2 \ar[rr]_{\bar{g}_2}&& G_3 }\]
	As usual, this last switch brings identities $\check{m}_0=\bar{f}_1\circ u_1$ and $g_2\circ j_0=\bar{g}_2\circ u_0$. Applying the second point of \Cref{lem:indep-global-left} to $\dder{D}_0\cdot S_{j_0, j_1}(\dder{D}_2)\cdot S_{j_0, j_1}(\dder{D}_1)$ and the fact that $(\X, \R)$ is well-switching we get the following diagrams

@@

We next prove some technical lemma that will be useful to show the
possibility of transforming a derivation into a switch equivalent one
by switching steps only along inversions.

@@

\begin{lemma}
  \label{le:switchB}
  Let $\der{D}_i$ for $i \in \interval[0]{n}$ be derivation sequences
  and consider a switching sequence consisting only of inversions
  \begin{center}
    $\der{D}_0 \shift{\nu_1} \der{D}_1 \shift{\nu_2} \ldots
    \shift{\nu_n} \der{D}_n$
  \end{center}
  Let $i \in \interval{n}$ be such that
  $\transp{i} \in \inv{\nu_{1,n}}$ and assume that steps $i$ and $i+1$ in $\der{D}_0$ are sequentially independent. Then there is a switching sequence
  consisting only of inversions that starts with $\transp{i}$, i.e.,
  \begin{center}
    $\der{D}_0 \shift{\transp{i}} \der{D}_1' \shift{\nu_2'} \ldots
    \shift{\nu_n'} \der{D}_n$
  \end{center}
\end{lemma}
